Towards Patient-Specific Neural Field Modeling using Micro-Electrode Data

### 2011-02-12 ltx/data_paper/DataPaper.tex
edits to maximization

Since the state equation is non-linear,  calculation of covariance and  cross-covariance terms required for the maximisation step is not straightforward as in the linear case. Therefore, an additional  Taylor expansion step to the URTSS algorithm is also created to approximate $q(\mathbf{x}_t)$ at each time instant which is to be used in the M-step. This is accomplished by taking the first-order truncated Taylor-series expansion of the sigmoidal activation function  around the smoothed state, $\hat{\mathbf x}_t^{b}$ at each time instant, i.e.
\begin{equation}\label{eq:SigmoidTaylor}
 f\left(\boldsymbol \phi^\top\left(\mathbf r'\right)\mathbf x_t\right) \approx f\left(\boldsymbol \phi^\top(\mathbf r')\hat{\mathbf x}_t\right)+  \boldsymbol \phi^\top\left(\mathbf r'\right)\left(\mathbf x_t - \hat{\mathbf x}_t\right)f'\left(\boldsymbol \phi^\top\left(\mathbf r'\right)\hat{\mathbf x}_t\right)
\end{equation}
where 
\begin{equation}\label{eq:SigmoidDerivative}
 f'\left(\boldsymbol \phi^\top\left(\mathbf r'\right)\hat{\mathbf x}_t\right)=\varsigma f\left(\boldsymbol \phi^\top\left(\mathbf r'\right)\hat{\mathbf x}_t\right)\left( 1-f\left( \boldsymbol \phi^\top\left(\mathbf r'\right)\hat{\mathbf x}_t\right)\right)
\end{equation}
substituting \ref{eq:SigmoidTaylor} in \ref{eq:QmatrixForSigmapoints}  we get an approximation for $q(\mathbf x_t)$ around the smoothed estimate $\hat{\mathbf{x}}_t^b$ 
\begin{equation}\label{eq:qTaylor}
 q(\mathbf x_t) \approx q\left(\hat{\mathbf x}_t\right)+\int_\Omega \boldsymbol{\Psi}\left(\mathbf{r}'\right)\boldsymbol \phi^\top\left(\mathbf r'\right) \left(\mathbf x_t -\hat{\mathbf x}_t\right)f'\left(\boldsymbol \phi^\top\left(\mathbf r'\right)\hat{\mathbf x}_t\right) d\mathbf{r}'
\end{equation}
Define
\begin{equation}
	e(x) = q(x) - f(x) \nonumber
\end{equation}

@@

The goal of the M-step is to maximize a lower bound on the joint likelihood function, $p(\mathbf X,\mathbf Y;\boldsymbol \Theta)$, where $\mathbf X$ is the set of all states, $\mathbf Y$ is the set of all observations, and $\boldsymbol \Theta$ denotes the set of all unknown parameters, i.e.,

@@

	\Theta = [\theta ...]
 connectivity kernel parameters $\boldsymbol\theta$, disturbance variance $\sigma_d$ and the observation noise variance $\sigma_{\epsilon}$. We choose to maximise \dean{the joint}, $p(\mathbf X,\mathbf Y;\boldsymbol \Theta)$, since it is not  possible to maximise the likelihood function $p(\mathbf Y;\boldsymbol\Theta)$ directly. Increasing the joint likelihood function increases the likelihood function $p(\mathbf Y;\boldsymbol \Theta)$ \cite{McLachlan1997}. We proceed by expanding the joint distribution $p(\mathbf X,\mathbf Y;\boldsymbol \Theta)$

@@

 \mathcal Q(\boldsymbol \Theta,\boldsymbol\Theta')&=\mathbf E_{\boldsymbol\Theta'}\left[\sum_{t=0}^{T-1}2\ln p(\mathbf y_{t+1}|\mathbf x_{t+1};\sigma_{\epsilon})+\sum_{t=0}^{T-1}2\ln p(\mathbf x_{t+1}|\mathbf x_{t};\boldsymbol \theta, \sigma_d)+\sum_{t=0}^{T-1}2\ln p(\mathbf x_0)\right],

@@

 It should be noted that the last term is not a function of parameter set and can be ignored therefore $\mathcal Q$-function can be rewritten as

@@

 p\left(\mathbf y_{t+1}|\mathbf x_{t+1};\sigma_{\epsilon}\right)&= \mid\boldsymbol\Sigma_{\epsilon}\mid^{-\frac{1}{2}}  \dean{exp}^{-\frac{1}{2}(\mathbf y_{t+1}-\mathbf C\mathbf  x_{t+1})^\top\boldsymbol\Sigma_{\epsilon}^{-1}(\mathbf y_{t+1}-\mathbf C\mathbf  x_{t+1}) } \\
p(\mathbf x_{t+1}|\mathbf x_{t};\boldsymbol \theta ,\sigma_d)&= \mid\boldsymbol\Sigma_{e}\mid^{-\frac{1}{2}} \times e^{-\frac{1}{2}(\mathbf x_{t+1}-q(\mathbf  x_t)\boldsymbol\theta-\xi  \mathbf x_t)^\top\boldsymbol\Sigma_e^{-1}(\mathbf x_{t+1}-q( \mathbf x_t)\boldsymbol\theta-\xi \mathbf  x_t)}

@@

 2 \ln p(\mathbf x_{t+1}|\mathbf x_{t};\boldsymbol \theta ,\sigma_d) &=-n_x\ln\sigma_d^2-\frac{1}{\sigma_d^2}\mathrm{tr}\left\lbrace\mathbf x_{t+1} \mathbf x_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace+\frac{2}{\sigma_d^2}\mathbf x_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}q( \mathbf x_t)\boldsymbol\theta+\frac{2\xi}{\sigma_s^2} \mathrm{tr} \left\lbrace \mathbf x_t\mathbf x_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace \nonumber \\

@@

Having approximated $q\left(\mathbf x_t\right)$ in the E-step, by substituting  \ref{eq:Qfunctiontrace1} and \ref{eq:Qfunctiontrace2} into \ref{eq:QIntermsofJointDist} and taking the expectation of the log likelihood function for all time instants gives an approximate to the lower-bound that needs to be maximised to gain optimal parameter estimates. Therefore $\mathcal Q$-function can be approximated as  
\begin{eqnarray}\label{eq:Voldermont}
 \mathcal Q(\boldsymbol \Theta,\boldsymbol\Theta')&\approx& -(T-1)n_y\ln \sigma_{\epsilon}^2-\frac{1}{\sigma_{\epsilon}^2}\mathrm{tr}\left\lbrace\boldsymbol\sum_{t=0}^{T-1}\left[ (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1}) (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1})^\top+\mathbf C \mathbf P_{t+1}\mathbf C^\top\right] \right\rbrace\nonumber \\
&-&(T-1)n_x\ln\sigma_d^2-\frac{1}{\sigma_d^2}\mathrm{tr}\left\lbrace \Xi_{0} \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace +\frac{2\xi}{\sigma_d^2} \mathrm{tr}\left\lbrace \Xi_{1} \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace+2\frac{1}{\sigma_d^2}\Xi_{2}\boldsymbol\theta
&-&\frac{1}{\sigma_d^2}\boldsymbol\theta^\top \Xi_{3}\boldsymbol\theta-\frac{2\xi}{\sigma_d^2} \Xi_{4} \boldsymbol\theta -  \frac{\xi^2}{\sigma_d^2}\mathrm{tr} \left\lbrace\Xi_{5}\tilde{\boldsymbol\Sigma}_e^{-1} \right\rbrace 
\end{eqnarray}
	\Xi_{0}&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_{t+1}\mathbf x_{t+1}^\top\right]=\sum_{t=0}^{T-1}\left(\mathbf P_{t+1}+\mathbf{\hat x}_{t+1}\mathbf{\hat x}_{t+1}^\top\right)\label{eq:defofXi0} \\
\Xi_{1}&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_t\mathbf x_{t+1}^\top\right]=\sum_{t=0}^{T-1}\left(\mathbf M_{t+1}+\mathbf{\hat x}_t\mathbf{\hat x}_{t+1}^\top\right) \label{eq:defofXi1} \\
\Xi_{2}&=\sum_{t=0}^{T-1}\left[ \mathbf{\hat x}_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf{\hat x}_t)+\int_\Omega\boldsymbol \phi^\top(\mathbf r') \mathbf M_{t+1} \tilde{\boldsymbol\Sigma}_e^{-1} \boldsymbol{\Psi}(\mathbf{r}') f'(\boldsymbol \phi^\top(\mathbf r')\mathbf {\hat x}_t) d\mathbf{r}'\right] \\	
\Xi_{3}&=\sum_{t=0}^{T-1}\left[q^\top(\mathbf{\hat x}_t)\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf{\hat x}_t)+\Lambda_t(\mathbf r)\mathbf P_t \tilde{\Lambda}_t(\mathbf r)\right] \\
 \Xi_{4}&=\sum_{t=0}^{T-1}\left[ \mathbf{\hat x}_{t}^\top\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf{\hat x}_t)+\int_\Omega\boldsymbol \phi^\top(\mathbf r') \mathbf P_t \tilde{\boldsymbol\Sigma}_e^{-1}  \boldsymbol{\Psi}(\mathbf{r}') f'(\boldsymbol \phi^\top(\mathbf r')\mathbf {\hat x}_t) d\mathbf{r}'\right] \\
 \Xi_{5}&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_t\mathbf x_{t}^\top\right]=\sum_{t=0}^{T-1}\left(\mathbf P_t+\mathbf{\hat x}_t\mathbf{\hat x}_t^\top\right)
and
	 \Lambda_t(\mathbf r)&=\begin{bmatrix}\boldsymbol \lambda_t^{(1)}(\mathbf r) & \boldsymbol \lambda_t^{(2)}(\mathbf r)& \dots &\boldsymbol \lambda_t^{(i)}(\mathbf r)\end{bmatrix} \label{eq:Lambda}\\
	 \boldsymbol\lambda_t^{(i)}(\mathbf r)&=\int_{\Omega} \boldsymbol \Psi^\top(\mathbf r')\tilde{\boldsymbol\Sigma}_e^{-1}\phi_i(\mathbf r')f'(\boldsymbol\phi^\top(\mathbf r')\mathbf {\hat x}_t)d\mathbf r'\label{eq:lambdai}\\
	\tilde{\Lambda}_t(\mathbf r)&=\begin{bmatrix}\tilde{\boldsymbol \lambda}_t^{(1)}(\mathbf r) \\ \tilde{\boldsymbol \lambda}_t^{(2)}(\mathbf r) \\ \vdots \\ \tilde{\boldsymbol \lambda}_t^{(i)}(\mathbf r)\end{bmatrix} \label{eq:Lambdatilde}\\
	 \tilde{\boldsymbol \lambda}_t^{(i)}(\mathbf r)&=\int_{\Omega} \boldsymbol \Psi(\mathbf r')\phi_i(\mathbf r')f'(\boldsymbol\phi^\top(\mathbf r')\mathbf {\hat x}_t)d\mathbf r' \label{eq:lambdatildei}
It should be noted that the second term in equation \ref{eq:Voldermont} is obtained by adding and subtracting $\mathbf C \mathbf{\hat{x}}_{t+1} \mathbf{\hat{x}}_{t+1}^\top \mathbf C^\top$ to the expectation, $\mathbf E_{\boldsymbol\Theta'}\left[ .\right] $. Given the state estimates, $\mathbf {\hat x^{b}}_t$ the state covariance $\mathbf P_t^b$ and cross-covariance $\mathbf M_t^b$ , the $\Xi$-variables can be computed directly. In the M-step, an update $\boldsymbol\Theta'$ is chosen such that $\mathcal{Q}\left(\boldsymbol\Theta',\boldsymbol\Theta'\right)>\mathcal{Q}\left(\boldsymbol\Theta,\boldsymbol\Theta'\right)$, this will guarantee that the likelihood increases.

Consider first the maximization of $\mathcal{Q}$-function with respect to $\boldsymbol\theta$, we can write
\begin{equation}\label{eq:Qpartialtheta}
 \frac{\partial \mathcal Q}{\partial \boldsymbol\theta}=\frac{2}{\sigma_d^2}\Xi_2-\frac{1}{\sigma_d^2}\boldsymbol\theta^\top(\Xi_3^\top+\Xi_3)-\frac{2\xi}{\sigma_d^2}\Xi_4
\end{equation}
noting that $\Xi_3$ is symmetric and equating \ref{eq:Qpartialtheta} to the zero vector we get
\begin{equation}\label{eq:theta}
 \boldsymbol \theta=\Xi_3^{-1}\left(\Xi_2^\top-\xi\Xi_4^\top \right).
\end{equation}
The second derivative is negative definite at this stationary point, and hence it is a global maximizer of $\mathcal{Q}$-function.
Now maximizing with respect to $\sigma_{\epsilon}^2$ we get
\begin{equation}
  \frac{\partial \mathcal Q}{\partial \sigma_{\epsilon}^2}=-(T-1)n_y\frac{1}{\sigma_{\epsilon}^2}+\frac{1}{\sigma_{\epsilon}^4}\mathrm{tr}\left\lbrace\sum_{t=0}^{T-1}\left[ (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1}) (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1})^\top+\mathbf C \mathbf P_{t+1}\mathbf C^\top\right] \right\rbrace
\end{equation}
setting above to zero and solving for $\sigma_{\epsilon}^2$ we have
\begin{equation}
 \sigma_{\epsilon}^2=\frac{1}{(T-1)n_y}\sum_{t=0}^{T-1}\mathrm{tr}\left\lbrace (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1}) (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1})^\top+\mathbf C \mathbf P_{t+1}\mathbf C^\top \right\rbrace
\end{equation}
A second derivative test shows that this is indeed the maximum. Finally differentiating with respec to $\sigma_d^2$ we get
\begin{eqnarray}
 \frac{\partial \mathcal Q}{\partial \sigma_d^2}&=& -(T-1)n_x\frac{1}{\sigma_d^2}+\frac{1}{\sigma_d^4}\mathrm{tr}\left\lbrace \Xi_0 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace-\frac{2\xi}{\sigma_d^4} \mathrm{tr}\left\lbrace \Xi_1 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace -2\frac{1}{\sigma_d^4}\Xi_2\boldsymbol\theta \nonumber \\
&+&\frac{1}{\sigma_d^4}\boldsymbol\theta^\top \Xi_3\boldsymbol\theta+\frac{2\xi}{\sigma_d^4} \Xi_4 \boldsymbol\theta +\frac{\xi^2}{\sigma_d^4}\mathrm{tr} \left\lbrace\Xi_5\tilde{\boldsymbol\Sigma}_e^{-1} \right\rbrace 
\end{eqnarray}
setting above to zero and solving for $\sigma_e^2$ we have
\begin{eqnarray}
\sigma_d^2=\frac{1}{(T-1)n_x}\left[ \mathrm{tr}\left\lbrace \Xi_0 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace-
2\xi\mathrm{tr}\left\lbrace \Xi_1 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace -2\Xi_2\boldsymbol\theta+\boldsymbol\theta^\top \Xi_3\boldsymbol\theta+2\xi\Xi_4 \boldsymbol\theta +\xi^2\mathrm{tr} \left\lbrace\Xi_5\tilde{\boldsymbol\Sigma}_e^{-1} \right\rbrace \right]  
\end{eqnarray}
The second derivative test confirms this is a maximum.

\section{Results}
\begin{itemize}
	\item Will show parameter distributions
	\item Will show parameter changes over time through the seizure
\end{itemize}
\section{M-Step-II}
\dean{Dean to check expansion.} Now we rearrange the equation so we can write down (or isolate (reason for doing this step is..)) the cross-covariance etc so we can use the information from the filter. Taking trace and rearranging, using the invariant cyclic permutations property of the trace, this distribution can be written as
\begin{align}\label{eq:Qfunctiontrace1}
  2\ln p\left(\mathbf y_{t+1}|\mathbf x_{t+1};\sigma_{\epsilon}\right) &= -n_y\ln \sigma_{\epsilon}^2-\frac{1}{\sigma_{\epsilon}^2}\mathrm{tr}\left\lbrace(\mathbf y_{t+1}-\mathbf C\mathbf  x_{t+1}) (\mathbf y_{t+1}-\mathbf C\mathbf  x_{t+1})^\top\right\rbrace    \\
 2 \ln p(\mathbf x_{t+1}|\mathbf x_{t};\boldsymbol \theta ,\sigma_d) &=-n_x\ln\sigma_d^2-\frac{1}{\sigma_d^2}\mathrm{tr}\left\lbrace\mathbf x_{t+1} \mathbf x_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace+\frac{2}{\sigma_d^2}\mathbf x_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}q( \mathbf x_t)\boldsymbol\theta+\frac{2\xi}{\sigma_d^2} \mathrm{tr} \left\lbrace \mathbf x_t\mathbf x_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace \nonumber \\
&-\frac{1}{\sigma_d^2}\boldsymbol\theta^\top q^\top(\mathbf x_t)\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf x_t)\boldsymbol\theta-\frac{2\xi}{\sigma_d^2} \mathbf x_t^\top\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf x_t)\boldsymbol\theta-\frac{\xi^2}{\sigma_d^2}\mathrm{tr}\left\lbrace \mathbf x_t \mathbf x_t^\top\tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace. \label{eq:Qfunctiontrace2}
\end{align}
Note that we ignore $\ln\mid\tilde{\boldsymbol\Sigma}_e\mid $ as it is not a function of the parameter set.

@@

	\Xi_{0}&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_{t+1}\mathbf x_{t+1}^\top\right]=\sum_{t=0}^{T-1}\left(\mathbf P_{t+1}+\mathbf{\hat x}_{t+1}\mathbf{\hat x}_{t+1}^\top\right)\label{eq:defofXi0} \\
\Xi_{1}&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_t\mathbf x_{t+1}^\top\right]=\sum_{t=0}^{T-1}\left(\mathbf M_{t+1}+\mathbf{\hat x}_t\mathbf{\hat x}_{t+1}^\top\right) \label{eq:defofXi1} \\
\Xi_{2}&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_t\mathbf x_{t}^\top\right]=\sum_{t=0}^{T-1}\left(\mathbf P_t+\mathbf{\hat x}_t\mathbf{\hat x}_t^\top\right)\\
\Xi_{3}&=E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}q( \mathbf x_t)\right] \\	
\Xi_{4}&= \mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}  q^\top(\mathbf  x_t)\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf x_t)\right]\\
 \Xi_{5}&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1} \mathbf x_t^\top\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf x_t)  \right]

@@

 \frac{\partial \mathcal Q}{\partial \boldsymbol\theta}=\frac{2}{\sigma_d^2}\Xi_2-\frac{1}{\sigma_d^2}\boldsymbol\theta^\top(\Xi_3^\top+\Xi_3)-\frac{2\xi}{\sigma_d^2}\Xi_4

@@

noting that $\Xi_3$ is symmetric and equating \ref{eq:Qpartialtheta} to the zero vector we get

@@

 \boldsymbol \theta=\Xi_3^{-1}\left(\Xi_2^\top-\xi\Xi_4^\top \right).

@@

\begin{eqnarray}
 \frac{\partial \mathcal Q}{\partial \sigma_d^2}&=& -(T-1)n_x\frac{1}{\sigma_d^2}+\frac{1}{\sigma_d^4}\mathrm{tr}\left\lbrace \Xi_0 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace-\frac{2\xi}{\sigma_d^4} \mathrm{tr}\left\lbrace \Xi_1 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace -2\frac{1}{\sigma_d^4}\Xi_2\boldsymbol\theta \nonumber \\
&+&\frac{1}{\sigma_d^4}\boldsymbol\theta^\top \Xi_3\boldsymbol\theta+\frac{2\xi}{\sigma_d^4} \Xi_4 \boldsymbol\theta +\frac{\xi^2}{\sigma_d^4}\mathrm{tr} \left\lbrace\Xi_5\tilde{\boldsymbol\Sigma}_e^{-1} \right\rbrace 
\end{eqnarray}

@@

\begin{eqnarray}

@@

2\xi\mathrm{tr}\left\lbrace \Xi_1 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace -2\Xi_2\boldsymbol\theta+\boldsymbol\theta^\top \Xi_3\boldsymbol\theta+2\xi\Xi_4 \boldsymbol\theta +\xi^2\mathrm{tr} \left\lbrace\Xi_5\tilde{\boldsymbol\Sigma}_e^{-1} \right\rbrace \right]  
\end{eqnarray}

>>>>

@@

The goal of the M-step is to maximize a lower bound on the joint likelihood function, $p(\mathbf X,\mathbf Y;\boldsymbol \Theta)$, where $\mathbf X$ is the set of all states, $\mathbf Y$ is the set of all observations, and $\boldsymbol \Theta$ denotes the set of all unknown parameters, i.e., connectivity kernel parameters $\boldsymbol\theta$, disturbance variance $\sigma_d$ and the observation noise variance $\sigma_{\epsilon}$.

@@

	\boldsymbol\Theta = \left\lbrace \boldsymbol\theta, \sigma_d, \sigma_{\epsilon}\right\rbrace 
\end{equation} We choose to maximise \dean{the joint}, $p(\mathbf X,\mathbf Y;\boldsymbol \Theta)$, since it is not  possible to maximise the likelihood function $p(\mathbf Y;\boldsymbol\Theta)$ directly. Increasing the joint likelihood function increases the likelihood function $p(\mathbf Y;\boldsymbol \Theta)$ \cite{McLachlan1997}. We proceed by expanding the joint distribution $p(\mathbf X,\mathbf Y;\boldsymbol \Theta)$

@@

 \mathcal Q(\boldsymbol \Theta,\boldsymbol\Theta')&=\mathbf E_{\boldsymbol\Theta'}\left[\sum_{t=0}^{T-1}2\ln p(\mathbf y_{t+1}|\mathbf x_{t+1};\sigma_{\epsilon})+\sum_{t=0}^{T-1}2\ln p(\mathbf x_{t+1}|\mathbf x_{t};\boldsymbol \theta. \sigma_d)+\sum_{t=0}^{T-1}2\ln p(\mathbf x_0)\right].

@@

 It should be noted that the last term is not a function of parameter set and can be ignored, therefore $\mathcal Q$-function can be rewritten as

@@

 p\left(\mathbf y_{t+1}|\mathbf x_{t+1};\sigma_{\epsilon}\right)&= \mid\boldsymbol\Sigma_{\epsilon}\mid^{-\frac{1}{2}}  \mathrm{exp}\left({-\frac{1}{2}\left(\mathbf y_{t+1}-\mathbf C\mathbf  x_{t+1}\right)^\top\boldsymbol\Sigma_{\epsilon}^{-1}\left(\mathbf y_{t+1}-\mathbf C\mathbf  x_{t+1}\right)}\right)\\
p\left(\mathbf x_{t+1}|\mathbf x_{t};\boldsymbol \theta ,\sigma_d\right)&= \mid\boldsymbol\Sigma_{e}\mid^{-\frac{1}{2}}\mathrm{exp}\left(-\frac{1}{2}\left(\mathbf x_{t+1}-q\left(\mathbf  x_t\right)\boldsymbol\theta-\xi  \mathbf x_t\right)^\top\boldsymbol\Sigma_e^{-1}\left(\mathbf x_{t+1}-q\left( \mathbf x_t\right)\boldsymbol\theta-\xi \mathbf  x_t\right)\right)

@@

 2 \ln p(\mathbf x_{t+1}|\mathbf x_{t};\boldsymbol \theta ,\sigma_d) &=-n_x\ln\sigma_d^2-\frac{1}{\sigma_d^2}\mathrm{tr}\left\lbrace\mathbf x_{t+1} \mathbf x_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace+\frac{2}{\sigma_d^2}\mathbf x_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}q( \mathbf x_t)\boldsymbol\theta+\frac{2\xi}{\sigma_d^2} \mathrm{tr} \left\lbrace \mathbf x_t\mathbf x_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace \nonumber \\

@@

@@

	\Xi_{0}&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_{t+1}\mathbf x_{t+1}^\top\right]\label{eq:defofXi0} \\
\Xi_{1}&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_t\mathbf x_{t+1}^\top\right]\label{eq:defofXi1} \\
\Xi_{2}&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_t\mathbf x_{t}^\top\right]\\
\Xi_{3}&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}q( \mathbf x_t)\right] \label{eq:Xi3} \\	
\Xi_{4}&= \mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}  q^\top(\mathbf  x_t)\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf x_t)\right] \label{eq:Xi4}\\
 \Xi_{5}&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1} \mathbf x_t^\top\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf x_t)  \right] \label{eq:Xi5}

@@

 \frac{\partial \mathcal Q}{\partial \boldsymbol\theta}=\frac{2}{\sigma_d^2}\Xi_3-\frac{1}{\sigma_d^2}\boldsymbol\theta^\top(\Xi_4^\top+\Xi_4)-\frac{2\xi}{\sigma_d^2}\Xi_5

@@

noting that $\Xi_4$ is symmetric and positive-definite by construction, equating \ref{eq:Qpartialtheta} to the zero vector we get

@@

 \boldsymbol \theta=\Xi_4^{-1}\left(\Xi_3^\top-\xi\Xi_5^\top \right).

@@

 \frac{\partial \mathcal Q}{\partial \sigma_d^2}&= -(T-1)n_x\frac{1}{\sigma_d^2}+\frac{1}{\sigma_d^4}\mathrm{tr}\left\lbrace \Xi_0 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace-\frac{2\xi}{\sigma_d^4} \mathrm{tr}\left\lbrace \Xi_1 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace+\frac{\xi^2}{\sigma_d^4}\mathrm{tr} \left\lbrace\Xi_2\tilde{\boldsymbol\Sigma}_e^{-1} \right\rbrace  \nonumber \\
&-\frac{2}{\sigma_d^4}\Xi_3\boldsymbol\theta +\frac{1}{\sigma_d^4}\boldsymbol\theta^\top \Xi_4\boldsymbol\theta+\frac{2\xi}{\sigma_d^4} \Xi_5 \boldsymbol\theta 

@@

@@

2\xi\mathrm{tr}\left\lbrace \Xi_1 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace +\xi^2\mathrm{tr} \left\lbrace\Xi_2\tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace -2\Xi_3\boldsymbol\theta+\boldsymbol\theta^\top \Xi_4\boldsymbol\theta+2\xi\Xi_5 \boldsymbol\theta\right]  
